Proofs for ``Geometry of Local Spectral Clustering''

### 2021-01-31 jmlr/revision/archive/proofs.tex
Final draft to be submitted to JMLR uploaded, in revision/ppr_clustering_neighborhood_graph

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{xr}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

@@

\paragraph{$\infty$-transportation distance between empirical and population measures.}
We now review the relevant upper bound on $\Delta_{\infty}(\Pbb,\Pbb_n)$, which holds under the following mild regularity conditions.
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:bounded_density} 
	The distribution $\Pbb$ has density $g: D \to (0,\infty)$ such that there exist $g_{\min} \leq 1 \leq g_{\max}$ for which
	\begin{equation*}
	(\forall x \in D)~~ g_{\min} \leq g(x) \leq g_{\max}.
	\end{equation*}
	\item 
	\label{asmp:domain} 
	The distribution $\Pbb$ is defined on a bounded, connected, open domain $D \subseteq \Rd$. If $d \geq 2$ then additionally $D$ has Lipschitz boundary.
\end{enumerate}
When $d = 1$, it follows from Proposition~6.2 of \textcolor{red}{Dudley} that $\Delta_{\infty}(\Pbb,\Pbb_n) \leq B_4 \|F - F_n\|_{\infty}$ for some positive constant $B_4$, and in turn from the DKW inequality (\textcolor{red}{Massart}) that
\begin{equation}
\label{eqn:ot_distance_1d}
\Delta_{\infty}(\Pbb,\Pbb_n) \leq B_4 \sqrt{\frac{\ln(2n/B_2)}{n}}
\end{equation}
with probability at least $1 - B_2/n$.

When $d \geq 2$, \cite{garciatrillos16b} derive an upper bound on the transportation distance $\Delta_{\infty}(\mbb{P},\mbb{P}_n)$.
\begin{theorem}[Theorem~1.1 of \cite{garciatrillos16}]
	\label{thm:garciatrillos16}
	Suppose $\Pbb$ satisfies~\ref{asmp:domain} and~\ref{asmp:bounded_density}. Then, there exists positive constants $B_{2}$ and $B_4$ that do not depend on $n$, such that with probability at least $1 - B_2/n$:
	\begin{equation*}
	\Delta_{\infty}(\mbb{P},\mbb{P}_n) \leq B_4 \cdot 
	\begin{dcases*}
	\frac{\ln(n)^{3/4}}{n^{1/2}},&~~\textrm{if $d = 2$,} \\
	\frac{\ln(n)^{1/d}}{n^{1/d}},&~~\textrm{if $d \geq 3$.}
	\end{dcases*}
	\end{equation*}
\end{theorem}
Assuming the candidate cluster $\mc{C}$ and conditional distribution $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}, then~\eqref{eqn:ot_distance_1d} ($d = 1)$ or Theorem~\ref{thm:garciatrillos16} ($d \geq 2$) apply to $\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n)$; we will use these upper bounds on $\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n)$ to show~\eqref{eqn:sample_to_population_conductance}.

@@

	\Psi_r(\mbb{Q}_2) \geq \Psi_r(\mbb{Q}_1) \cdot \biggl(1 - \frac{2b_4\Delta_T(\mbb{Q}_1,\mbb{Q}_2)}{\bigl(1 - \Psi_r(\mbb{Q}_1)\bigr) \cdot \bigl(d_{\min}(\mbb{Q}_2)\bigr)^2}\biggr) - \frac{b_4 \Delta_T(\mbb{Q}_1,\mbb{Q}_2)}{\bigl(1 - \Psi_r(\mbb{Q}_1)\bigr) \cdot \bigl(d_{\min}(\mbb{Q}_2)\bigr)^2},

@@

	where $b_4 := 2d \nu_d r^{d - 1} \cdot \max_{x \in \Rd}\{g(x)\}$ is a positive constant that does not depend on $\mbb{Q}_2$. 

@@

	R_{r - 2\Delta_{12} ,\mbb{Q}_1}^{(2\Delta_{12})}(x) \leq \nu_dr^d\Bigl\{1 - \Bigl(1 - \frac{2\Delta_{12}}{r}\Bigr)^d \Bigr\} \cdot \max_{x \in \Rd}\{g(x)\} \leq \underbrace{2 d \nu_d r^{d - 1} \cdot \max_{x \in \Rd}\{g(x)\}}_{= b_4} \cdot \Delta_{12},

@@

	R_{r,\mbb{Q}_1}^{(2\Delta_{12})}(x) \leq \nu_dr^d\Bigl\{\Bigl(1 + \frac{2\Delta_{12}}{r}\Bigr)^d - 1\Bigr\} \cdot \max_{x \in \Rd}\{g(x)\} \leq 2 b_4 \cdot \Delta_{12}.

@@

	\Phi_{\mbb{Q}_2,r}(\mc{S}) & \geq \frac{\cut_{\mbb{Q}_1,r}\bigl(T^{-1}(\mc{S})\bigr) - b_4 \Delta_{12}}{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) + 2b_4 \Delta_{12}} \\ & = \Phi_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) \cdot \biggl(\frac{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S}))}{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) + 2b_4\Delta_{12}}\biggr) - \frac{b_4\Delta_{12}}{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) + 2b_4\Delta_{12}} \\
	& \overset{\eqref{pf:conductance_lb_transportation_distance_2}}{\geq} \Phi_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) \cdot \biggl(\frac{\wb{\vol}_{\mbb{Q}_2,r}(\mc{S}) - 2b_4\Delta_{12}}{\wb{\vol}_{\mbb{Q}_2,r}(\mc{S})}\biggr) - \frac{b_4\Delta_{12}}{\wb{\vol}_{\mbb{Q}_2,r}(\mc{S})}.

@@

	\inf_{\mc{S}: \mc{S} \in \mathfrak{L}_{r}(\mbb{Q}_1,\mbb{Q}_2)} \Phi_{\mbb{Q}_2,r}(\mc{S}) \geq \Psi_{r}(\mbb{Q}_1) \cdot \biggl(1 - \frac{2 b_4 \Delta_{12}}{(1 - \Psi_r(\mbb{Q}_1)) \cdot d_{\min}(\mbb{Q}_2)^2}\biggr) - \frac{b_4\Delta_{12}}{(1 - \Psi_r(\mbb{Q}_1)) \cdot d_{\min}(\mbb{Q}_2)^2}

@@

\paragraph{Putting the pieces together.} 
First, we note that
\begin{equation*}
\Psi_r(\wt{\mbb{P}}) = \Psi_{\wt{\mbb{P}},r}(\mathrm{supp}(\wt{\mbb{P}})) = \Psi_{\Pbb,r}(\mc{C}), ~~\textrm{and}~~\Psi_{r}(\wt{\mbb{P}}_n) = \Psi_{\wt{\mbb{P}}_n, r}(\mathrm{supp}(\wt{\mbb{P}}_n)) = \Psi_{n,r}(\mc{C}[X]),
\end{equation*}
so that we may apply Proposition~\ref{prop:conductance_lb_transportation_distance} to get a lower bound on $\Psi_{n,r}(\mc{C}[X])$ in terms of $\Psi_{\Pbb,r}(\mc{C})$, $\Delta_{\infty}(\wt{\Pbb}, \wt{\Pbb}_n)$, and $d_{\min}(\wt{\Pbb}_{n})$. We now use the bounds we have derived on transportation distance and minimum degree. From the derivations in Section~\ref{subsec:sample_to_population_conductance}, it follows with probability at least $1 - (n + 1)\exp\{-nb_2/16\}$ that,
\begin{equation*}
d_{\min}(\wt{\Pbb}_n) = \frac{1}{\wt{n}} \bigl(d_{\min}(\wt{G}_{n,r}) + 1\bigr) \geq \frac{1}{\sqrt{2}} d_{\min}(\wt{\Pbb}).
\end{equation*}
% Here I replaced a factor of (1 - 1/sqrt{2})^2 by 1/16, for convenience.
On the other hand, taking
\begin{equation*}
b_5 := \frac{1}{2b_4}\Psi_r(\wt{\Pbb}) \cdot (1 - \Psi_r(\wt{\Pbb})) \cdot d_{\min}(\wt{\Pbb})^2,~~\textrm{and}~~ B_1 := \frac{1}{B_4}\min\Bigl\{\frac{b_5}{2\Psi_r(\wt{\Pbb})}, b_5, \frac{r}{4(d - 1)} \Bigr\},
\end{equation*}
by~\eqref{eqn:ot_distance_1d} (if $d = 1$) or Theorem~\ref{thm:garciatrillos16} (if $d \geq 2$) along with~\eqref{eqn:sample_to_population_conductance_sample_complexity}, we have that
\begin{equation*}
\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n) \leq B_4 \frac{(\log n)^{p_d}}{\min\{n^{1/2},n^{1/d}\}} \leq \min\Bigl\{\frac{b_5}{2\Psi_r(\wt{\Pbb})}, b_5, \frac{r}{4(d - 1)} \Bigr\} \cdot \delta
\end{equation*}
with probability at least $1 - B_2/n$. Therefore by Proposition~\ref{prop:conductance_lb_transportation_distance},
\begin{equation*}
\Psi_r(\wt{\Pbb}_n) \geq \Psi_r(\wt{\Pbb}) \cdot \biggl(1 - \frac{2b_4\Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb})}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2}\biggr) - \frac{b_4 \Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb}_N)}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2} \geq \Psi_r(\wt{\Pbb}) (1 - 2\delta)
\end{equation*}
with probability at least $1 - B_2/n - (n + 1)\exp\{-nb_2/16\}$, establishing~\eqref{eqn:sample_to_population_conductance} upon taking $b_3 := b_2/16$.

@@

\begin{proof}[Proof of~\eqref{eqn:local_spread_ub}.]
	\textcolor{red}{(TODO)}
\end{proof}

@@

\subsection{Proof of Theorem~\ref{thm:ppr_lb}}
We start by defining some constants, to make our proof statements easier to digest. Put
\begin{align*}
C_{3,\delta} & := \frac{72(1 + \delta)}{(1 - \delta)} \sqrt{8/3 + 8 \delta}
,&&C_{4,\delta} := \frac{72}{(1 - 3\delta)\pi}, \\
B_{1,\delta} & := 768 \cdot (1 + 3\delta) \cdot \ln\Bigl(C_{4,\delta} \frac{\rho\sigma}{r^2 \epsilon^2} \Bigr),&&B_{2,\delta} := \frac{(1 + \delta)^2}{(1 - \delta)^2} \cdot \frac{\rho \sigma}{r^2 \epsilon^2} \\
B_4 &:= \frac{6 \sigma \rho}{(\sigma - 2r)(\rho - r)},&&B_5 := 1 + \frac{12\sigma\rho}{r^2} + \frac{2\rho}{r} \\
b_6 &:=  b_7 \wedge \cut_{\Pbb,r}(\mc{L}) \wedge d_{\min}(\Pbb)/14 \wedge \vol_{\Pbb,r}(\mc{L} \cap \mc{C}^{(1)}),&&b_7 := \vol_{\Pbb,r}(\mc{X})/4 \wedge \frac{\epsilon r^2}{4\rho\sigma} \wedge \frac{\pi r^3}{8\sigma \rho^2}.
\end{align*}
To prove Theorem~\ref{thm:ppr_lb}, we use Theorem~\ref{thm:normalized_cut_ppr}, Proposition~\ref{prop:sample_to_population_1} and~\eqref{eqn:ncut_lower_set} to show that the cluster estimate $\wh{C}$ must have a small normalized cut. On the other hand, in Lemma~\ref{lem:normalized_cut_lb} we establish that any set $Z \subseteq X$ which is close to $\mc{C}^{(1)}[X]$---meaning $\vol_{n,r}(Z \vartriangle \mc{C}^{(1)}[X])$ is small---has a large normalized cut.
\begin{lemma}
	\label{lem:normalized_cut_lb}
	Fix $\delta \in (0,1)$. With probability at least $1 - B_5\exp\{-n\delta^2b_7\}$, the following statement holds:
	\begin{equation}
	\label{eqn:normalized_cut_lb}
	\Phi_{n,r}(Z) \geq \frac{(1 - \delta)^2}{2(1 + \delta)\pi} \left(1 - 2 \frac{\sigma \rho}{(1 - \delta) r^2 n^2} \vol_{n,r}(Z \vartriangle \mc{C}^{(1)}[X]) \right) \frac{\epsilon^2 r}{\sigma},~~\textrm{for all $Z \subseteq X$}.
	\end{equation}
\end{lemma}
We therefore conclude that $\vol_{n,r}(\wh{C} \vartriangle \mc{C}^{(1)}[X])$ must be large. In the remainder of this, we detail the probabilistic estimates used in the proof of Theorem~\ref{thm:ppr_lb}, and then give a formal proof of Theorem~\ref{thm:ppr_lb} and then of Lemma~\ref{lem:normalized_cut_lb}.

\textit{Probabilistic estimates.}
In addition to~\eqref{eqn:normalized_cut_lb}, we will assume (i) that the graph normalized cut of $\mc{L}$ and local spread of $\mc{X}$ are respectively upper and lower bounded,
\begin{equation*}
\Phi_{n,r}(\mc{L}[X]) \leq (1 + 3\delta) \cdot \Phi_{\Pbb,r}(\mc{L}),~~\textrm{and}~~s_{n,r}(X) \geq (1 - 3\delta) \cdot s_{\Pbb,r}(\mc{X});
\end{equation*}
(ii) that the graph volume of $\mc{L}$ is upper and lower bounded,
\begin{equation*}
(1 - \delta) \vol_{\Pbb,r}(\mc{L}) \leq \frac{1}{n(n - 1)}\vol_{n,r}(\mc{L}[X]) \leq (1 + \delta) \vol_{\Pbb,r}(\mc{L});
\end{equation*}
(iii) that the graph volumes of $\mc{L} \cap \mc{C}^{(1)}$ and $\mc{C}^{(1)}$ are respectively lower and upper bounded,
\begin{equation}
(1 - \delta) \vol_{\Pbb,r}(\mc{L} \cap \mc{C}^{(1)}) \leq \frac{1}{n(n - 1)}\vol_{n,r}({\mc{L}[X] \cap \mc{C}^{(1)}}[X])~~\textrm{and}~~\frac{1}{n(n - 1)}\vol_{n,r}(\mc{C}^{(1)}[X]) \leq (1 + \delta) \vol_{\Pbb,r}(\mc{C}^{(1)});
\end{equation}
(iv) that the graph volume of $\mc{X}$ is lower bounded,
\begin{equation*}
\frac{1}{n(n - 1)}\vol_{n,r}(X) \geq (1 - \delta) \vol_{\Pbb,r}(\mc{X});
\end{equation*}
and finally (v) that the maximum degree of $G_{n,r}$ is upper bounded,
\begin{equation*}
\frac{1}{n - 1}d_{\max}(G_{n,r}) \leq (1 + \delta) d_{\max}(\Pbb).
\end{equation*}
It follows from Lemma~\ref{lem:normalized_cut_lb}, Propositions~\ref{prop:sample_to_population_1} and \ref{prop:sample_to_population_2}, and Lemmas~\ref{lem:hoeffding_2} and~\ref{lem:bernstein_union} that these estimates are together satisfied with probability at least $1 - B_5\exp\{-n\delta^2b_7\} - 3\exp\{-n\delta^2 \cut_{\Pbb,r}(\mc{L})\} - (2n + 2)\exp\{-n\delta^2 \cdot d_{\min}(\Pbb)/14\} - 5\exp\{-n\delta^2 \vol_{\Pbb,r}(\mc{L} \cap \mc{C}^{(1)})\} \geq 1 - (B_5 + 2n + 10)\exp\{-n\delta^2 b_6\}$.

@@

\section{Experimental Details}
\label{sec:experimental_details}

\textcolor{red}{(TODO)}

>>>>

\documentclass{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage{xr}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{graphics}
\usepackage[export]{adjustbox} 
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

@@

\paragraph{$\infty$-transportation distance between empirical and population measures.}
We now review the relevant upper bound on $\Delta_{\infty}(\Pbb,\Pbb_n)$, which holds under the following mild regularity conditions.
\begin{enumerate}[label=(A\arabic*)]
	\item 
	\label{asmp:bounded_density} 
	The distribution $\Pbb$ has density $g: D \to (0,\infty)$ such that there exist $g_{\min} \leq 1 \leq g_{\max}$ for which
	\begin{equation*}
	(\forall x \in D)~~ g_{\min} \leq g(x) \leq g_{\max}.
	\end{equation*}
	\item 
	\label{asmp:domain} 
	The distribution $\Pbb$ is defined on a bounded, connected, open domain $D \subseteq \Rd$. If $d \geq 2$ then additionally $D$ has Lipschitz boundary.
\end{enumerate}
When $d = 1$, it follows from Proposition~6.2 of \textcolor{red}{Dudley} that $\Delta_{\infty}(\Pbb,\Pbb_n) \leq B_5 \|F - F_n\|_{\infty}$ for some positive constant $B_5$, and in turn from the DKW inequality (\textcolor{red}{Massart}) that
\begin{equation}
\label{eqn:ot_distance_1d}
\Delta_{\infty}(\Pbb,\Pbb_n) \leq B_5 \sqrt{\frac{\ln(2n/B_2)}{n}}
\end{equation}
with probability at least $1 - B_2/n$.

When $d \geq 2$, \cite{garciatrillos16b} derive an upper bound on the transportation distance $\Delta_{\infty}(\mbb{P},\mbb{P}_n)$.
\begin{theorem}[Theorem~1.1 of \cite{garciatrillos16}]
	\label{thm:garciatrillos16}
	Suppose $\Pbb$ satisfies~\ref{asmp:domain} and~\ref{asmp:bounded_density}. Then, there exists positive constants $B_{2}$ and $B_5$ that do not depend on $n$, such that with probability at least $1 - B_2/n$:
	\begin{equation*}
	\Delta_{\infty}(\mbb{P},\mbb{P}_n) \leq B_5 \cdot 
	\begin{dcases*}
	\frac{\ln(n)^{3/4}}{n^{1/2}},&~~\textrm{if $d = 2$,} \\
	\frac{\ln(n)^{1/d}}{n^{1/d}},&~~\textrm{if $d \geq 3$.}
	\end{dcases*}
	\end{equation*}
\end{theorem}
Assuming the candidate cluster $\mc{C}$ and conditional distribution $\wt{\Pbb}$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_density}, then~\eqref{eqn:ot_distance_1d} ($d = 1)$ or Theorem~\ref{thm:garciatrillos16} ($d \geq 2$) apply to $\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n)$; we will use these upper bounds on $\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n)$ to show~\eqref{eqn:sample_to_population_conductance}.

@@

	\Psi_r(\mbb{Q}_2) \geq \Psi_r(\mbb{Q}_1) \cdot \biggl(1 - \frac{2B_6\Delta_T(\mbb{Q}_1,\mbb{Q}_2)}{\bigl(1 - \Psi_r(\mbb{Q}_1)\bigr) \cdot \bigl(d_{\min}(\mbb{Q}_2)\bigr)^2}\biggr) - \frac{B_6 \Delta_T(\mbb{Q}_1,\mbb{Q}_2)}{\bigl(1 - \Psi_r(\mbb{Q}_1)\bigr) \cdot \bigl(d_{\min}(\mbb{Q}_2)\bigr)^2},

@@

	where $B_6 := 2d \nu_d r^{d - 1} \cdot \max_{x \in \Rd}\{g(x)\}$ is a positive constant that does not depend on $\mbb{Q}_2$. 

@@

	R_{r - 2\Delta_{12} ,\mbb{Q}_1}^{(2\Delta_{12})}(x) \leq \nu_dr^d\Bigl\{1 - \Bigl(1 - \frac{2\Delta_{12}}{r}\Bigr)^d \Bigr\} \cdot \max_{x \in \Rd}\{g(x)\} \leq \underbrace{2 d \nu_d r^{d - 1} \cdot \max_{x \in \Rd}\{g(x)\}}_{= B_6} \cdot \Delta_{12},

@@

	R_{r,\mbb{Q}_1}^{(2\Delta_{12})}(x) \leq \nu_dr^d\Bigl\{\Bigl(1 + \frac{2\Delta_{12}}{r}\Bigr)^d - 1\Bigr\} \cdot \max_{x \in \Rd}\{g(x)\} \leq 2 B_6 \cdot \Delta_{12}.

@@

	\Phi_{\mbb{Q}_2,r}(\mc{S}) & \geq \frac{\cut_{\mbb{Q}_1,r}\bigl(T^{-1}(\mc{S})\bigr) - B_6 \Delta_{12}}{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) + 2B_6 \Delta_{12}} \\ & = \Phi_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) \cdot \biggl(\frac{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S}))}{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) + 2B_6\Delta_{12}}\biggr) - \frac{B_6\Delta_{12}}{\wb{\vol}_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) + 2B_6\Delta_{12}} \\
	& \overset{\eqref{pf:conductance_lb_transportation_distance_2}}{\geq} \Phi_{\mbb{Q}_1,r}(T^{-1}(\mc{S})) \cdot \biggl(\frac{\wb{\vol}_{\mbb{Q}_2,r}(\mc{S}) - 2B_6\Delta_{12}}{\wb{\vol}_{\mbb{Q}_2,r}(\mc{S})}\biggr) - \frac{B_6\Delta_{12}}{\wb{\vol}_{\mbb{Q}_2,r}(\mc{S})}.

@@

	\inf_{\mc{S}: \mc{S} \in \mathfrak{L}_{r}(\mbb{Q}_1,\mbb{Q}_2)} \Phi_{\mbb{Q}_2,r}(\mc{S}) \geq \Psi_{r}(\mbb{Q}_1) \cdot \biggl(1 - \frac{2 B_6 \Delta_{12}}{(1 - \Psi_r(\mbb{Q}_1)) \cdot d_{\min}(\mbb{Q}_2)^2}\biggr) - \frac{B_6\Delta_{12}}{(1 - \Psi_r(\mbb{Q}_1)) \cdot d_{\min}(\mbb{Q}_2)^2}

@@

\paragraph{Putting the pieces together.} 
First, we note that
\begin{equation*}
\Psi_r(\wt{\mbb{P}}) = \Psi_{\wt{\mbb{P}},r}(\mathrm{supp}(\wt{\mbb{P}})) = \Psi_{\Pbb,r}(\mc{C}), ~~\textrm{and}~~\Psi_{r}(\wt{\mbb{P}}_n) = \Psi_{\wt{\mbb{P}}_n, r}(\mathrm{supp}(\wt{\mbb{P}}_n)) = \Psi_{n,r}(\mc{C}[X]),
\end{equation*}
so that we may apply Proposition~\ref{prop:conductance_lb_transportation_distance} to get a lower bound on $\Psi_{n,r}(\mc{C}[X])$ in terms of $\Psi_{\Pbb,r}(\mc{C})$, $\Delta_{\infty}(\wt{\Pbb}, \wt{\Pbb}_n)$, and $d_{\min}(\wt{\Pbb}_{n})$. We now use the bounds we have derived on transportation distance and minimum degree. From the derivations in Section~\ref{subsec:sample_to_population_conductance}, it follows with probability at least $1 - (n + 1)\exp\{-nb_2/16\}$ that,
\begin{equation*}
d_{\min}(\wt{\Pbb}_n) = \frac{1}{\wt{n}} \bigl(d_{\min}(\wt{G}_{n,r}) + 1\bigr) \geq \frac{1}{\sqrt{2}} d_{\min}(\wt{\Pbb}).
\end{equation*}
% Here I replaced a factor of (1 - 1/sqrt{2})^2 by 1/16, for convenience.
On the other hand, taking
\begin{equation*}
b_6 := \frac{1}{2B_6}\Psi_r(\wt{\Pbb}) \cdot (1 - \Psi_r(\wt{\Pbb})) \cdot d_{\min}(\wt{\Pbb})^2,~~\textrm{and}~~ B_1 := B_5 \Bigl(\min\Bigl\{\frac{b_6}{2\Psi_r(\wt{\Pbb})}, b_6, \frac{r}{4(d - 1)} \Bigr\}\Bigr)^{-1},
\end{equation*}
by~\eqref{eqn:ot_distance_1d} (if $d = 1$) or Theorem~\ref{thm:garciatrillos16} (if $d \geq 2$) along with~\eqref{eqn:sample_to_population_conductance_sample_complexity}, we have that
\begin{equation*}
\Delta_{\infty}(\wt{\Pbb},\wt{\Pbb}_n) \leq B_5 \frac{(\log n)^{p_d}}{\min\{n^{1/2},n^{1/d}\}} \leq \min\Bigl\{\frac{b_6}{2\Psi_r(\wt{\Pbb})}, b_6, \frac{r}{4(d - 1)} \Bigr\} \cdot \delta
\end{equation*}
with probability at least $1 - B_2/n$. Therefore by Proposition~\ref{prop:conductance_lb_transportation_distance},
\begin{equation*}
\Psi_r(\wt{\Pbb}_n) \geq \Psi_r(\wt{\Pbb}) \cdot \biggl(1 - \frac{2B_6\Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb})}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2}\biggr) - \frac{B_6 \Delta_{\infty}(\wt{\Pbb}_n,\wt{\Pbb}_N)}{\bigl(1 - \Psi_r(\wt{\Pbb})\bigr) \cdot \bigl(d_{\min}(\wt{\Pbb}_n)\bigr)^2} \geq \Psi_r(\wt{\Pbb}) (1 - 2\delta)
\end{equation*}
with probability at least $1 - B_2/n - (n + 1)\exp\{-nb_2/16\}$, establishing~\eqref{eqn:sample_to_population_conductance} upon taking $b_3 := b_2/16$.

@@

\begin{proof}[Proof of~\eqref{eqn:local_spread_ub}.]
	The statement follows since 
	\begin{equation*}
	d_{\min}(\Pbb) \geq \frac{\pi r^2}{2} \cdot \min_{x \in \mc{X}} f(x) = \frac{\pi r^2}{2} \cdot \frac{\epsilon}{\rho \sigma}
	\end{equation*}
	and
	\begin{equation*}
	\vol_{\Pbb,r}(\mc{X}) \leq d_{\max}(\Pbb) \leq \frac{\pi r^2}{2 \sigma \rho}.
	\end{equation*}
\end{proof}

@@

\subsection{Proof of Theorem~\ref{thm:ppr_lb}}
We start by defining some constants, to make our proof statements easier to digest. Put
\begin{align*}
C_{3,\delta} & := \frac{72(1 + \delta)}{(1 - \delta)} \sqrt{8/3 + 8 \delta}
,&&C_{4,\delta} := \frac{72}{(1 - 3\delta)\pi}, \\
B_{1,\delta} & := 768 \cdot (1 + 3\delta) \cdot \ln\Bigl(C_{4,\delta} \frac{\rho\sigma}{r^2 \epsilon^2} \Bigr),&&B_{2,\delta} := \frac{(1 + \delta)^2}{(1 - \delta)^2} \cdot \frac{\rho \sigma}{r^2 \epsilon^2} \\
B_4 & := 1 + \frac{12\sigma\rho}{r^2} + \frac{2\rho}{r},&&b_4 :=  b_8 \wedge \cut_{\Pbb,r}(\mc{L}) \wedge d_{\min}(\Pbb)/14 \wedge \vol_{\Pbb,r}(\mc{L} \cap \mc{C}^{(1)}), \\
b_8 &:= \vol_{\Pbb,r}(\mc{X})/4 \wedge \frac{\epsilon r^2}{4\rho\sigma} \wedge \frac{\pi r^3}{8\sigma \rho^2}.
\end{align*}
To prove Theorem~\ref{thm:ppr_lb}, we use Theorem~\ref{thm:normalized_cut_ppr}, Proposition~\ref{prop:sample_to_population_1} and~\eqref{eqn:ncut_lower_set} to show that the cluster estimate $\wh{C}$ must have a small normalized cut. On the other hand, in Lemma~\ref{lem:normalized_cut_lb} we establish that any set $Z \subseteq X$ which is close to $\mc{C}^{(1)}[X]$---meaning $\vol_{n,r}(Z \vartriangle \mc{C}^{(1)}[X])$ is small---has a large normalized cut.
\begin{lemma}
	\label{lem:normalized_cut_lb}
	Fix $\delta \in (0,1)$. With probability at least $1 - B_4\exp\{-n\delta^2b_8\}$, the following statement holds:
	\begin{equation}
	\label{eqn:normalized_cut_lb}
	\Phi_{n,r}(Z) \geq \frac{(1 - \delta)^2}{2(1 + \delta)\pi} \left(1 - 2 \frac{\sigma \rho}{(1 - \delta) r^2 n^2} \vol_{n,r}(Z \vartriangle \mc{C}^{(1)}[X]) \right) \frac{\epsilon^2 r}{\sigma},~~\textrm{for all $Z \subseteq X$}.
	\end{equation}
\end{lemma}
We therefore conclude that $\vol_{n,r}(\wh{C} \vartriangle \mc{C}^{(1)}[X])$ must be large. In the remainder of this, we detail the probabilistic estimates used in the proof of Theorem~\ref{thm:ppr_lb}, and then give a formal proof of Theorem~\ref{thm:ppr_lb} and then of Lemma~\ref{lem:normalized_cut_lb}.

\textit{Probabilistic estimates.}
In addition to~\eqref{eqn:normalized_cut_lb}, we will assume (i) that the graph normalized cut of $\mc{L}$ and local spread of $\mc{X}$ are respectively upper and lower bounded,
\begin{equation*}
\Phi_{n,r}(\mc{L}[X]) \leq (1 + 3\delta) \cdot \Phi_{\Pbb,r}(\mc{L}),~~\textrm{and}~~s_{n,r}(X) \geq (1 - 3\delta) \cdot s_{\Pbb,r}(\mc{X});
\end{equation*}
(ii) that the graph volume of $\mc{L}$ is upper and lower bounded,
\begin{equation*}
(1 - \delta) \vol_{\Pbb,r}(\mc{L}) \leq \frac{1}{n(n - 1)}\vol_{n,r}(\mc{L}[X]) \leq (1 + \delta) \vol_{\Pbb,r}(\mc{L});
\end{equation*}
(iii) that the graph volumes of $\mc{L} \cap \mc{C}^{(1)}$ and $\mc{C}^{(1)}$ are respectively lower and upper bounded,
\begin{equation}
(1 - \delta) \vol_{\Pbb,r}(\mc{L} \cap \mc{C}^{(1)}) \leq \frac{1}{n(n - 1)}\vol_{n,r}({\mc{L}[X] \cap \mc{C}^{(1)}}[X])~~\textrm{and}~~\frac{1}{n(n - 1)}\vol_{n,r}(\mc{C}^{(1)}[X]) \leq (1 + \delta) \vol_{\Pbb,r}(\mc{C}^{(1)});
\end{equation}
(iv) that the graph volume of $\mc{X}$ is lower bounded,
\begin{equation*}
\frac{1}{n(n - 1)}\vol_{n,r}(X) \geq (1 - \delta) \vol_{\Pbb,r}(\mc{X});
\end{equation*}
and finally (v) that the maximum degree of $G_{n,r}$ is upper bounded,
\begin{equation*}
\frac{1}{n - 1}d_{\max}(G_{n,r}) \leq (1 + \delta) d_{\max}(\Pbb).
\end{equation*}
It follows from Lemma~\ref{lem:normalized_cut_lb}, Propositions~\ref{prop:sample_to_population_1} and \ref{prop:sample_to_population_2}, and Lemmas~\ref{lem:hoeffding_2} and~\ref{lem:bernstein_union} that these estimates are together satisfied with probability at least $1 - B_4\exp\{-n\delta^2b_8\} - 3\exp\{-n\delta^2 \cut_{\Pbb,r}(\mc{L})\} - (2n + 2)\exp\{-n\delta^2 \cdot d_{\min}(\Pbb)/14\} - 5\exp\{-n\delta^2 \vol_{\Pbb,r}(\mc{L} \cap \mc{C}^{(1)})\} \geq 1 - (B_4 + 2n + 10)\exp\{-n\delta^2 b_4\}$.

@@

\section{Experimental Details}
\label{sec:experimental_details}
Finally, we detail the settings of our experiments, and include an additional figure.
